Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation

### 2022-03-17 Writing/appendix.rmd
some texts added for explaining why yield prediction does not mean good eonr prediction

```{r ct-ex, fig.width=5, fig.height=5, fig.align="center", fig.cap = "Figure A.1: An illustrative example of a causal tree"}
par(family="Times")
fancyRpartPlot(opfit, sub="")
```

@@

In the first node, all the samples ($1024$ observations) belong to the node, and the causal effect of the treatment was calculated as the difference in the mean yield of the treated and control groups, which turned out to be $593$ kg/ha in this example. The first data split criterion is whether the $\textit{ecs}$ value is greater or less than $25$. The observations with $\textit{ecs}$ greater than or equal to $25$ move to the left node, while those with $\textit{ecs}$ less than $25$ move to the right. For each of the subsamples, the treatment effect is identified as the difference in the mean yields of the treated and control groups within the subsample. The treatment effect for the low-$\textit{ecs}$ observations was $619$ kg/ha, and for the high-$\textit{ecs}$ observations was $982$ kg/ha. This means that for the impact of the increase in seed rate the treatment effect is heterogeneous over space, having a larger positive impact on yield in high-$\textit{ecs}$ areas. Trees are further developed to have more groups of subsamples to allow for more flexible representations of the heterogeneous treatment effects. The estimated heterogeneous treatments effects can then be used to identify site-specific EOIRs. It is noteworth that at no point CF estimates yield "level."

>>>>

```{r ct-ex, fig.width=5, fig.height=5, fig.align="center", fig.cap = "Figure A.1: An illustrative example of a causal tree"}
par(family = "Times")
fancyRpartPlot(opfit, sub = "")
```

@@

In the first node, all the samples ($1024$ observations) belong to the node, and the causal effect of the treatment was calculated as the difference in the mean yield of the treated and control groups, which turned out to be $593$ kg/ha in this example. The first data split criterion is whether the $\textit{ecs}$ value is greater or less than $25$. The observations with $\textit{ecs}$ greater than or equal to $25$ move to the left node, while those with $\textit{ecs}$ less than $25$ move to the right. For each of the subsamples, the treatment effect is identified as the difference in the mean yields of the treated and control groups within the subsample. The treatment effect for the high-$\textit{ecs}$ observations was $619$ kg/ha, and for the low-$\textit{ecs}$ observations was $982$ kg/ha. This means that for the impact of the increase in seed rate the treatment effect is heterogeneous over space, having a larger positive impact on yield in low-$\textit{ecs}$ areas. Trees are further developed to have more groups of subsamples to allow for more flexible representations of the heterogeneous treatment effects. The estimated heterogeneous treatments effects can then be used to identify site-specific EOIRs. It is noteworth that at no point CF estimates yield "level."

### 2022-03-17 Writing/manuscript_cea.rmd
some texts added for explaining why yield prediction does not mean good eonr prediction

Interestingly, CNN completely failed in its estimations of site-specific EONRs. This happened because the site-specific yield response functions identified by CNN were linear, and all with the same slope. Therefore, CNN failed to capture a declining marginal product of N. The constant marginal product caused the estimated EONR to take on either the lowest or highest experimental N application rate. To examine whether this problem is a product solely of the specific choice of CNN architecture, we also tested another CNN model, which was analogous to CNN-ST proposed by @barbosa2020modeling. 
This architecture also estimated that yield and N rates were perfectly linearly related, and so estimated site-specific EONR poorly. Note that @barbosa2020modeling just compared models through the lens of yield prediction but did not use the estimated models to derive EONRs. In their work, in some cases, CNN predicted yield well; but in all cases predicted EONR poorly.

Figure \@ref(fig:dist-optN) and Figure \@ref(fig:dist-piLoss) show the distributions of the one thousand RMSEs of the EONR estimations and $\hat{\pi}_{def}$ by ML method and modeling scenario, respectively. The center of distribution mass of the CF-base's RMSE values and $\hat{\pi}_{def}$ are to the left of the centers of the distribution masses of the other methods' RMSE values, showing that CF-base more accurately and consistently estimated EONR across all the modeling scenarios. 
Interestingly, RF seems particularly vulnerable to the inclusion of more variables, with RMSE ($\hat{\pi}_{def}$) increasing from `r table_optN_prep[type=="test" & Model=="aby",rmse_optN_RF]` (`r table_optN_prep[type=="test" & Model=="aby", pi_loss_RF]`) in the "aby" scenario to `r table_optN_prep[type=="test" & Model=="aabbyytt", rmse_optN_RF]` (`r table_optN_prep[type=="test" & Model=="aabbyytt", pi_loss_RF]`) in the "aabbyytt" scenario. CF-base was more robust to the inclusion of additional variables. Overall, the simulations present CF-base as the clear winner in estimating EONR.

@@

For RF, BRF and CNN, Figure \@ref(fig:plot-y-optN) plots the RMSE values of EONR estimation against the RMSE values of yield prediction, and the red lines show the best-fit line describing the relationship between them. For all the models the regression line was almost flat, showing little if any correlation between the two measures. In other words, the quality of a model's yield predictions provided little if any information about the quality of its EONR predictions.

```{r plot-y-optN, fig.width=7, fig.height=7.5, fig.cap = "The relationship in the performance between EONR estimation and yield prediction, by ML method and modeling scenario"}
fig_y_optN
```

`r ftext('NOTE: The red line in the figures denotes the best-fit line to explain the relationship between the RMSEs of EONR prediction and yield prediction.', fp_text(font.size = 9, font.family = "Times New Roman"))`

Table \@ref(tab:table-y) shows the mean RMSE of yield prediction over the one thousand simulations. Since CF-base does not predict yield, the table only reports the results from RF, BRF and CNN. Compared to the result of EONR estimation, RF performed as well as BRF in the simple modeling scenarios of "aby" and "abytt." This fact along with the poor performance of RF in predicting EONR show that strong predictive power of yield levels does not imply accurate prediction of EONR. Further, RF shows vulnerability to the inclusion of many correlated variables, as is displayed for EONR prediction as well. While CNN performed worse than RF and BRF in "aby" and "abytt," it performed better than RF in more complicated modeling scenarios, showing robustness, at least in yield prediction, to inclusion of many correlated variables. Of course, while on average in the "aabbyy" and "aabbyytt" scenarios CNN predicted yields better than RF did, CNN's EONR predictions were worse than RF's, which further backs our claim that a good yield prediction capability does not imply good EONR prediction capability.

@@

@@

All previous research has followed a two-step procedure to estimate site-specific EONRs. The first step was to select the model that predicts yield most accurately, and the following step used the selected model to predict EONR. Table \@ref(tab:count-table) shows that following this process led to the choice of the best (as measured by profit) EONR prediction model in fewer than $30\%$ of the "aby" and "abytt" simulation rounds, $40\%$ of the "aabbyy" simulation rounds, and $50\%$ of the "aabbyytt" simulation rounds. For example, the table shows that, in the "aby" scenario, RF attained the lowest RMSE for yield prediction in `r summary_res_CNN_RF_BRF[Model=="aby", count_y_RF]` of 1000 iterations, and attained the lowest $\hat{\pi}_{def}$ in `r summary_res_CNN_RF_BRF[Model=="aby", count_RF]` of those `r summary_res_CNN_RF_BRF[Model=="aby", count_y_RF]` iterations. That is, in `r summary_res_CNN_RF_BRF[Model=="aby", count_y_RF]-summary_res_CNN_RF_BRF[Model=="aby", count_RF]` out of `r summary_res_CNN_RF_BRF[Model=="aby", count_y_RF]` rounds, the RF models best at yield prediction did not provide EONR estimates that led to high profits. Similarly, in no case did the CNN method provide both the best yield prediction and the best EONR estimation simultaneously. In the rounds in which BRF predicted yield best, the two-step process worked well. For example, Table \@ref(tab:count-table) shows that BRF predicted EONR best in `r summary_res_CNN_RF_BRF[Model=="aby", count_BRF]` of the `r summary_res_CNN_RF_BRF[Model=="aby", count_y_BRF]` "aby" simulation rounds in which it predicted yields best. But BRF predicted yields best in fewer than $30\%$ of the rounds, implying that the two-step procedure failed more than $70\%$ of the time.

>>>>

Interestingly, CNN completely failed in its estimations of site-specific EONRs. This happened because the site-specific yield response functions identified by CNN were linear, and all with the same slope. Therefore, CNN failed to capture a declining marginal product of N. The constant marginal product caused the estimated EONR to take on either the lowest or highest experimental N application rate. To examine whether this problem is a product solely of the specific choice of CNN architecture, we also tested another CNN model, which was analogous to CNN-ST proposed by @barbosa2020modeling. 
This architecture also estimated that yield and N rates were perfectly linearly related, and so estimated site-specific EONR poorly. Note that @barbosa2020modeling just compared models through the lens of yield prediction but did not use the estimated models to derive EONRs.

Figure \@ref(fig:dist-optN) and Figure \@ref(fig:dist-piLoss) show the distributions of the RMSEs of the EONR estimations and $\hat{\pi}_{def}$ obtained from the 1000 simualtion rounds by ML method and modeling scenario, respectively. The center of distribution mass of the CF-base's RMSE values and $\hat{\pi}_{def}$ are to the left of the centers of the distribution masses of the other methods' RMSE values, showing that CF-base more accurately and consistently estimated EONR across all the modeling scenarios. 
Interestingly, RF seems particularly vulnerable to the inclusion of more variables, with RMSE ($\hat{\pi}_{def}$) increasing from `r table_optN_prep[type=="test" & Model=="aby",rmse_optN_RF]` (`r table_optN_prep[type=="test" & Model=="aby", pi_loss_RF]`) in the "aby" scenario to `r table_optN_prep[type=="test" & Model=="aabbyytt", rmse_optN_RF]` (`r table_optN_prep[type=="test" & Model=="aabbyytt", pi_loss_RF]`) in the "aabbyytt" scenario. CF-base was more robust to the inclusion of additional variables. Overall, the simulations present CF-base as the clear winner in estimating EONR.

@@

Table \@ref(tab:table-y) shows the mean RMSE of yield prediction over the one thousand simulations. Since CF-base does not predict yield, the table only reports the results from RF, BRF and CNN. Compared to the result of EONR estimation, RF performed as well as BRF in the simple modeling scenarios of "aby" and "abytt." This fact along with the poor performance of RF in predicting EONR presented in \@ref(tab:table-optN) show that strong predictive power of yield levels does not imply accurate prediction of EONR. Further, RF shows vulnerability to the inclusion of many correlated variables, as is displayed for EONR prediction as well. While CNN performed worse than RF and BRF in "aby" and "abytt," it performed better than RF in more complicated modeling scenarios, showing robustness, at least in yield prediction, to inclusion of many correlated variables. Of course, while on average in the "aabbyy" and "aabbyytt" scenarios CNN predicted yields better than RF did, CNN's EONR predictions were worse than RF's, which further backs our claim that a good yield prediction capability does not imply good EONR prediction capability.

@@

Table \@ref(tab:xxx) illustrates the distinction between yield prediction and yield response function prediction.

In essence, good performance in predicinting yield "levels" at the observed points does not mean good performanc ein predicting yield response function, which is about finding out what "would" happen to yield if rates other than the one actually applied were applied. This happens because you can predict yield leves for wrong reasons by giving wrong amount of credit to wrong drivers (variables) of yield levels. This is because prediction-oriented ML methods focus on predicting yield levels well at the "observed" points, but not the points that were not seen. This is by construction because they are trained to minimize the error in yield "level" prediction at the "observed" points. In contrast, CF focuses to recover the yield response functions well (or alternatively the causal impact of N on yield at different discrete rates of N). Our results how that this distinction between prediction-oriented ML methods and CF can result in economically meanigful difference, where CF outperforms.

@@

All previous research has followed a two-step procedure to estimate site-specific EONRs. The first step was to select the model that predicts yield most accurately, and the following step used the selected model to predict EONR. Table \@ref(tab:count-table) shows that following this process led to the choice of the best (as measured by profit) EONR prediction model in fewer than $30\%$ of the "aby" and "abytt" simulation rounds, $40\%$ of the "aabbyy" simulation rounds, and $50\%$ of the "aabbyytt" simulation rounds. For example, the table shows that, in the "aby" scenario, RF attained the lowest RMSE for yield prediction in `r summary_res_CNN_RF_BRF[Model=="aby", count_y_RF]` of 1000 iterations, and attained the lowest $\hat{\pi}_{def}$ in `r summary_res_CNN_RF_BRF[Model=="aby", count_RF]` of those `r summary_res_CNN_RF_BRF[Model=="aby", count_y_RF]` iterations. That is, in `r summary_res_CNN_RF_BRF[Model=="aby", count_y_RF]-summary_res_CNN_RF_BRF[Model=="aby", count_RF]` out of `r summary_res_CNN_RF_BRF[Model=="aby", count_y_RF]` rounds, the RF model performed the best at yield prediction among RF, BRF, and CNN, but did not provide EONR estimates that led to the highest profit. Similarly, in no case did the CNN method provide both the best yield prediction and the best EONR estimation simultaneously. In the rounds in which BRF predicted yield best, the two-step process worked well. For example, Table \@ref(tab:count-table) shows that BRF predicted EONR best in `r summary_res_CNN_RF_BRF[Model=="aby", count_BRF]` of the `r summary_res_CNN_RF_BRF[Model=="aby", count_y_BRF]` "aby" simulation rounds in which it predicted yields best. But BRF predicted yields best in fewer than $30\%$ of the rounds, implying that the two-step procedure failed more than $70\%$ of the time.
